---
title: "analysis"
author: "Team 17: Lizbeth Ramirez, Shruthi Leon, Eric Tsai, and Andrew Kim"
format: revealjs
---

## 1. Introduction

This project analyzes factors influencing song popularity using Bayesian modeling.

## 2. Research Questions

This analysis explores three research questions:

1. What song features most strongly predict song popularity?
2. Has the relationship between song loudness and popularity changed over time?
3. How much of song popularity comes from artist effects?

## 3. Data Preparation

```{r echo=TRUE, results='hide',message=FALSE, warning=FALSE}
# load libraries
library(bayesrules)
library(tidyverse)
library(rstan)
library(bayesplot)
library(broom.mixed)

# load dataset
music <- read_csv("../data/music_clean.csv")
```
## 4. Analysis
## Question 1: What song features most strongly predict song popularity?
```{r}
# EDA
# Model
# Results 
```
#### Interpretation

## Question 2: Has the relationship between song loudness and popularity changed over time?
```{r}
# EDA
# Model
# Results 
```
#### Interpretation

## Question 3: How much of song popularity comes from artist effects?
```{r}
# EDA
# Model
# Results 
```
#### Interpretation

## 5. Conclusion


